This documentation is intended to get you started, but we understand
that SimWrapper is a bit different from other data viz platforms and so
you may have trouble getting started. And while the docs are extensive,
they are never ``complete'' and could always be clearer\ldots{}

\hypertarget{example-dashboards-and-configurations}{%
\subsection{Example dashboards and
configurations}\label{example-dashboards-and-configurations}}

When starting out, please look at examples of SimWrapper in action in
the
\href{https://github.com/simwrapper/simwrapper-example-project}{SimWrapper
example project}. If you clone that repo and point SimWrapper at it,
you'll see lots of examples of the various graphs and plugins available.

Then you can start modifying the configuration YAML files to experiment
with settings or to add your own data.

\hypertarget{asking-questions-and-making-suggestions}{%
\subsection{Asking questions and making
suggestions}\label{asking-questions-and-making-suggestions}}

The best place to ask a question is to file a
\href{https://github.com/simwrapper/simwrapper/issues}{GitHub issue
here}. That way, others can also see the question and eventual responses
and answers.

\begin{itemize}
\tightlist
\item
  Please label your questions with the label \texttt{Question} so we can
  better filter them from the long list of open bugs and issues.
\end{itemize}

\hypertarget{filing-bug-reports-issues}{%
\subsection{Filing bug reports /
issues}\label{filing-bug-reports-issues}}

We use GitHub Issues to track development, so that's the right place to
file bugs for SimWrapper, too.

\begin{itemize}
\tightlist
\item
  File a \href{https://github.com/simwrapper/simwrapper/issues}{GitHub
  issue here}.
\item
  Please label it as a \texttt{bug} thanks!
\end{itemize}

\hypertarget{contributing-via-github-pull-requests}{%
\subsection{Contributing via GitHub pull
requests}\label{contributing-via-github-pull-requests}}

We gladly accept PR's for documentation or the code itself!

\begin{itemize}
\tightlist
\item
  Code repository: \url{https://github.com/simwrapper/simwrapper}
\item
  Documentation repository:
  \url{https://github.com/simwrapper/simwrapper.github.io}
\end{itemize}
